# t-SNE analysis 

## Dataset Human
total 163 columns rpkm data  
kidney 20, 41, 42 miss 1 year  
kidney 29 miss pre  
kidney 19 miss 3 months  
```{r message=FALSE, include=FALSE}

library(data.table)
library(Rtsne)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
library(reshape2)
library(stringr)
library(monocle)
get_time_point <- function(x) substr(x, 18, nchar(x))
dt <- fread("C:/Users/Ryan/Documents/3.Kidney/HUMAN_all.gene_RPKM.xls")
## regexp: regexpr("[[:digit:]]_", names(dt)[-(1:7)], perl =T)
#table(substr(names(dt)[-(1:7)],1,16))
# total 163 columns rpkm data
# kidney 41, 20, 42 miss 1 year
# kidney 29 miss pre
# kidney 19 miss 3 months

setcolorder(dt,
            c(names(dt)[1:7],
              names(dt)[-(1:7)][order(names(dt)[-(1:7)])])
)

#dt[,ensembl_gene_id] rownames contains duplicates.
ge <- dt[, -(1:7)]
ge <- log2(ge + 1)
ge <- t(ge) # matrix

dummy <- rep(0, dim(ge)[2])
pre <- ge[grep("pre", rownames(ge)),]
post <- ge[grep("post", rownames(ge)),] ;post = post[-29,]
year <- ge[grep("year", rownames(ge)),] ;year = year[-29,]
year <- rbind(year[1:18,], dummy, year[19:38,], dummy, dummy)
month <- ge[grep("month", rownames(ge)),] ;month = month[-28,]
month <- rbind(month[1:18,], dummy, month[19:40,])


post.remove <- post - pre
year.remove <- (year - pre); year.remove <- year.remove[-grep("dummy",rownames(year.remove)),]
month.remove <- month - pre; month.remove <- month.remove[-grep("dummy",rownames(month.remove)),]

ge.remove <- rbind(post.remove, month.remove, year.remove )
Labels <- unname(sapply(rownames(ge.remove), get_time_point))
persons <- substr(rownames(ge.remove), 15, 16)
set.seed(100)
tsne <- Rtsne(ge.remove, dims = 2, perplexity = 30, verbose = T, max_iter = 500)
```

```{r human, fig.cap = "**Similarity between samples and time points(backgroud deducted).**"}
tbl <- tbl_df(tsne$Y)
tbl$persons = persons
tbl$time_point = Labels
tbl %>% ggplot(aes(x = V1, y = V2, label = persons, col = time_point)) +
  geom_text()

```
In figure \@ref(fig:human): __pre__ expression profile($log(FPKM + 1)$) was deducted from __post__, __1 year__, and __3 months__. And then the dimension reduction was done using [t-SNE](https://lvdmaaten.github.io/tsne/) algorithm.  
The plot makes more sense now: later time points(**3 months**, **1 year**) stick together and are separated from **post** points, upon which a *Pseudotime Sample Trajectory* was constructed by means of [Monocle 2](http://cole-trapnell-lab.github.io/monocle-release/docs/#constructing-single-cell-trajectories).  

## Dataset Mouse

```{r mouse, echo = F, fig.cap = 'Same t-SNE analysis for mouse data.'}
mouse <- fread("C:/Users/Ryan/Documents/3.Kidney/MOUSE_all.gene_RPKM.csv")

setnames(mouse, "V1", "EnsembleID")
mouse.ge <- mouse[, -(1:3)]
mouse.ge <- log2(mouse.ge + 1)
mouse.ge <- t(mouse.ge)

Labels <- colsplit(rownames(mouse.ge), pattern = "-", names = c("Mouse", "grp"))$Mouse
ind <- regexpr("[[:digit:]]", Labels, perl =T)
prefix = c(); time_point = c()
for(i in 1:length(Labels)) {
  prefix <- append(prefix, substr(Labels[i], 1, ind[i] - 1))
  time_point <- append(time_point, substr(Labels[i], ind[i], nchar(Labels[i])))
}

set.seed(81)
tsne <- Rtsne(mouse.ge, dims = 2, perplexity = 9, verbose = T, max_iter = 500)
tbl <- tbl_df(tsne$Y)
tbl$sample_type = prefix
tbl$time_point = time_point
tbl %>% ggplot(aes(x = V1, y = V2, label = time_point, col = sample_type)) + 
  geom_text()

```
As shown in the Figure \@ref(fig:mouse):  

1. The syntaxis of same time-point and same mouse-type are stronger in this Mouse data, though it has less sample points here--49 samples in total, divided into three types(IRI, NORM, SHAM), each time point with 3 replicate except for 6mN has 4.

2. The perplexity = 9 and set.seed(81) were tuned mannualy by hands-and-eyes. Based on tens of runs, the results are alike--the clustering is good. Perplexity >5 and <= 10 generally works here for limitied times of run done by myself.
